A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-15 documents/plm.tex
tweaks

The third general idea is that we need to control for over-fitting.
That is, once you have a flexible-enough model you can in principle fit \emph{anything},
  whether it was caused by the spacecraft, intrinsic stellar variability, or a transiting exoplanet.
How do you prevent a very flexible model from using small noise-induced fluctuations in all the input data
  from being carefully combined linearly into models for every nook and cranny in the target data?
In most projects in astrophysics, over-fitting is controlled for by limiting model freedom.
The model is restricted in the number of parameters
  (as in ``you can't have more parameters than data points'')
  or by limiting the dimensionality
  (as in principal components analysis)
  or by applying strong priors
  (as with smoothness priors, that effectively reduce freedom without explicitly reducing the number of parameters).
In each case, the restriction on the model can be set with tools like cross-validation,
  the fully marginalized likelihood (Bayes factor or evidence),
  chi-squared statistical criteria,
  or intuition or heuristics.
Here we take a different and more general approach, which is to use a train-and-test framework.

In this framework, the data used to \emph{train} the model%
  ---meaning set the values of the parameters of the model---%
  are disjoint from the \emph{test} data
  ---meaning the data that are being predicted (the target data).
Because we are extremely concerned with detecting Earth-analog exoplanet transits,
  which take about ten hours,
  we adopt an extreme version of the train-and-test framework,
  in which the training data are always separated from the test data by at least ten hours.
That is, when we are using the model to predict a particular pixel in the target data taken at time $t$,
  we use parameters obtained by an optimization that makes use of only data
  that comes either at times prior to $t-\Delta t$ or after $t+\Delta t$,
  where $\Delta t$ is a tunable parameter but which we will set to roughly $10$~h.
This ensures that no information about any exoplanet transit itself
  can be entering into the prediction of the pixels contributing to the stellar photometry.
In general, if there is a scientific goal of preserving intrinsic stellar variability,
  or transit signals,
  on time-scales of $\tau$,
  the parameter optimization ought to be based on training data taken with a time-exclusion zone of width $\Delta > \tau$.

@@

The fourth general idea involved in this kind of data-driven modeling is that the models aren't \emph{interpretable}.
This means, in particular, that although the model might do a good job of \emph{predicting} the target pixels,
  using a linear (or more complex) combination ofthe input pixel data,
  it won't deliver anything that can be unambiguously interpreted as the \emph{flux} of the star in question,
  or any other signal we care about.
The data-driven model \emph{effectively} describes the pointing, point-spread function, and flat-field
  of the telescope and camera,
  plus the variability of every star and every exoplanet transit,
  but it does so without ever \emph{explicitly} creating any of those objects.
We have to make some kind of heuristic or interpretive move to extract from the data-driven model the quanitites of interest.

@@

Finally, the fifth general idea is that there is no objective \emph{ground truth} against which we can tell
  that any particular data-driven model is better or worse than any other.
This problem is a problem for \name, and for the \Kepler\ PDC, and any other data-driven calibration models.
One might think that the ``best'' model is the one that predicts data with lowest variance;
  this would be true if all models adhered to the same train-and-test framework, which they don't.
Besides, a model that can predict not just the spacecraft-induced variability,
  but also variability caused by exoplanet transits of interest,
  will effectively over-fit and distort the most important information in the light-curves.
That is, what is considered best for modeling the data depends on the objectives of the user.
For us, who are interested (in the long term) in finding and characterizing Earth analogs,
  the ``best'' data-driven model is the one that produces the most success in finding and characterizing Earth analogs!
What we will show in what follows is that the \name\ photometry does not distort
  artificially exoplanet transit signals injected into real \Kepler\ pixel-level data.
We can also show that it produces stellar light-curves that look sensible for known variable stars.
We cannot yet show that it is optimal for our long-term goals,
  nor can we show that it produces images or light-curves that are more accurately calibrated;
  we will only show promising properties of the \name\ outputs.
In the end, the value of the \name\ will be demonstrated by the scientific projects it enables.

>>>>

The third general idea is that we need to control for over-fitting.
That is, once you have a flexible-enough model you can in principle fit \emph{anything},
  whether it was caused by the spacecraft, intrinsic stellar variability, or a transiting exoplanet.
How do you prevent a very flexible model from taking small noise-induced fluctuations in all the input data
  and carefully combining them linearly into detailed models for every nook and cranny in the target data?
In most projects in astrophysics, over-fitting is controlled for by limiting model freedom.
The model is restricted in the number of parameters
  (as in ``you can't have more parameters than data points'')
  or by limiting the dimensionality
  (as in principal components analysis)
  or by applying strong priors
  (as with smoothness priors, that effectively reduce freedom without explicitly reducing the number of parameters).
In each case, the restriction on the model is controlled by some \emph{hyper-parameter}
  (such as the number of inputs, or number of principal components, or the strength of the smoothness prior).
The hyper-parameters can be set or tested with tools like cross-validation,
  the fully marginalized likelihood (Bayes factor or evidence),
  chi-squared statistical criteria,
  or intuition or heuristics.
Here we take a different and more general approach, which is to use a train-and-test framework.

In this framework, the data used to \emph{train} the model%
  ---meaning set the values of the parameters of the model---%
  are disjoint from the \emph{test} data
  ---meaning the data that are being predicted (the target data).
Because we are extremely concerned with detecting Earth-analog exoplanet transits,
  which take about ten hours,
  we adopt an extreme version of the train-and-test framework,
  in which the training data are always separated from the test data by at least ten hours.
That is, when we are using the model to predict a particular pixel in the target data taken at time $t$,
  we use parameters obtained by an optimization that makes use of only data
  that comes either at times prior to $t-\Delta t$ or after $t+\Delta t$,
  where $\Delta t$ is a tunable parameter but which we will set to roughly $10$~h.
This ensures that no information about any exoplanet transit itself
  can be entering into the prediction of the pixels contributing to the stellar photometry.
In general, if there is a scientific goal of preserving intrinsic stellar variability,
  or transit signals,
  on time-scales of $\tau$,
  the parameter optimization ought to be based on training data taken with a time-exclusion zone of half-width $\Delta t > \tau$.

% need to choose consistent notation or terminology about the contiguous data chunks.

@@

The fourth general idea involved in this kind of data-driven modeling is that the models aren't \emph{interpretable}.
This means, in particular, that although the model might do a good job of \emph{predicting} the target pixels,
  using a linear (or more complex) combination of the input pixel data,
  it won't deliver anything that can be unambiguously interpreted as the \emph{flux} of the star in question,
  or any other signal we care about.
The data-driven model \emph{effectively} describes the pointing, point-spread function, and flat-field
  of the telescope and camera,
  plus the variability of every star and every exoplanet transit,
  but it does so without ever \emph{explicitly} creating any of those objects.
We have to make some kind of heuristic or interpretive move to extract from the data-driven model the quanitites of interest.

@@

Finally, the fifth general idea is that there is no objective \emph{ground truth} against which we can tell
  that any particular data-driven model is better or worse than any other.
This problem is a problem for \name, and for the \Kepler\ PDC, and any other data-driven calibration models.
One might think that the ``best'' model is the one that predicts data with lowest variance;
  this would be true if all models adhered to the same train-and-test framework, which they don't.
Besides, a model that can predict not just the spacecraft-induced variability,
  but also variability caused by exoplanet transits of interest,
  will effectively over-fit and distort the most important information in the light-curves.
That is, what is considered best for modeling the data depends on the objectives of the user.
For us, who are interested (in the long term) in finding and characterizing Earth analogs,
  the ``best'' data-driven model is the one that produces the most success in finding and characterizing Earth analogs!
What we will show in what follows is that the \name\ photometry does not distort
  artificial exoplanet transit signals injected into real \Kepler\ pixel-level data.
We can also show that it produces stellar light-curves that look sensible for known variable stars.
We cannot yet show that it is optimal for our long-term goals,
  nor can we show that it produces images or light-curves that are more accurately calibrated;
  we will only show promising properties of the \name\ outputs.
In the end, the value of the \name\ will be demonstrated by the scientific projects it enables.
